---
title: 'Google Case study: Bellabeat'
author: "Sen"
date: "2023-05-25"
output:
  pdf_document: default
  html_document: default
---

## Installing and loading common packages and libraries
```{r}
install.packages('tidyverse')
install.packages("dplyr")
install.packages("ggplot2")
install.packages("lubridate")
install.packages("janitor")
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(janitor)
```

## Loading the CSV files
```{r}
daily_activity <- read.csv("dailyActivity_merged.csv")
sleep_day <- read.csv("sleepDay_merged.csv")
calories_hourly <- read.csv("hourlyCalories_merged.csv")
intensities_hourly <- read.csv("hourlyIntensities_merged.csv")
weight_loginfo <- read.csv("weightLogInfo_merged.csv")
```

## Getting the unique participants each dataframe
```{r distinct users}
n_distinct(daily_activity$Id)
n_distinct(sleep_day$Id)
n_distinct(calories_hourly$Id)
n_distinct(intensities_hourly$Id)
n_distinct(weight_loginfo$Id)
```
## summarise data using summary function:

```{r activity info}
daily_activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes, Calories) %>%
  summary()
```

```{r active minutes per category}
daily_activity %>%
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes) %>%
  summary()
```

```{r calories info}
calories_hourly %>%
  select(Calories) %>%
  summary()
```

```{r sleep info}
sleep_day %>%
  select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>%
  summary()
```

```{r weight info}
weight_loginfo %>%
  select(WeightKg, BMI) %>%
  summary()
```


```{r Intensity info}
intensities_hourly %>%
  select(ActivityHour, TotalIntensity, AverageIntensity) %>%
  summary()
```

## Plotting a few explorations
Relationship between Total Steps and Sedentary Minutes: 
```{r}
ggplot(data=daily_activity, aes(x=TotalSteps, y=SedentaryMinutes)) + 
  geom_point() +
  geom_smooth() +
  labs(title="Total Steps vs. Sedentary Minutes")
```

Relationship between Total Steps and Calories?
```{r}
ggplot(data=daily_activity, aes(x=TotalSteps, y=Calories)) + 
  geom_point() + 
  geom_smooth() + 
  labs(title="Total Steps vs. Calories")
```

Relationship between minutes asleep and time in bed? You might expect it to be almost completely linear - are there any unexpected trends?
```{r}
ggplot(data=sleep_day, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) +
  geom_point() + 
  labs(title="Total Minutes Asleep vs. Total Time in Bed")
```

## Convert date time format.
```{r intensities}
intensities_hourly$ActivityHour=as.POSIXct(intensities_hourly$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
intensities_hourly$time <- format(intensities_hourly$ActivityHour, format = "%H:%M:%S")
intensities_hourly$date <- format(intensities_hourly$ActivityHour, format = "%m/%d/%y")
```

```{r activity}
daily_activity$ActivityDate=as.POSIXct(daily_activity$ActivityDate, format="%m/%d/%Y", tz=Sys.timezone())
daily_activity$date <- format(daily_activity$ActivityDate, format = "%m/%d/%y")
```

```{r sleep}
sleep_day$SleepDay=as.POSIXct(sleep_day$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
sleep_day$date <- format(sleep_day$SleepDay, format = "%m/%d/%y")
```


## Merging these two datasets together
```{r}
combined_data <- merge(sleep_day, daily_activity, by = c('Id', 'date'))
head(combined_data)
```

## Take a look at how many participants are in this data set.

```{r}
n_distinct(combined_data$Id)
```
## Relationship between Total Minutes Asleep and Sedentry Minutes.
```{r}
ggplot(data=combined_data, aes(x=TotalMinutesAsleep, y=SedentaryMinutes)) + 
geom_point(color='orange') + 
  geom_smooth() +
  labs(title="Minutes Asleep vs. Sedentary Minutes")
```


# aggregate data by day of week to summarize averages
```{r}
aggregate_data <- mutate(combined_data, day = wday(SleepDay, label = TRUE))
summary_activity_sleep_data <- aggregate_data %>% 
  group_by(day) %>% 
  summarise(AvgDailySteps = mean(TotalSteps),
            AvgAsleepMinutes = mean(TotalMinutesAsleep),
            AvgAwakeTimeInBed = mean(TotalTimeInBed), 
            AvgSedentaryMinutes = mean(SedentaryMinutes),
            AvgLightlyActiveMinutes = mean(LightlyActiveMinutes),
            AvgFairlyActiveMinutes = mean(FairlyActiveMinutes),
            AvgVeryActiveMinutes = mean(VeryActiveMinutes), 
            AvgCalories = mean(Calories))
head(summary_activity_sleep_data)
```

# Plot data by day of week to summarize averages
```{r} 
ggplot(data = summary_activity_sleep_data, mapping = aes(x = day, y = AvgDailySteps)) +
geom_col(fill = "orange") + labs(title = "Daily Step Count")
```

